How to Analyze Build Logs --- A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-24 tse-paper/content.tex
Integrate TODO

% \section{Introduction}
% TODO: Include term predictive CI, link to FB paper and
%https://www-sciencedirect-com.tudelft.idm.oclc.org/science/article/pii/S0164121218301730

@@

When a build or one of its steps fails, developers typically
scavenge the logs for information related to the source of the error.
This manual activity is
tedious and prone to error~\cite{santolucito2018statically}.
As a
single build log can be over 50 megabytes large~\cite{beller2017oops},
finding the small chunk of information in it which is linked to the error,
is akin to searching for a
needle in a haystack.
In addition to helping understand and fix build failures, the information
richness of build logs enables a plethora of other applications: with
the help of
build logs, we can better understand and categorize CI
failures and compile
errors~\cite{islam2017insights,seo2014programmers}, differentiate
testing practices~\cite{orellana2017differences,vassallo2017a-tale},
train classifiers to predict the build outcome and
duration~\cite{ni2017cost,bisong2017built,machalica2019predictive}, or
investigate static analysis tools in CI
builds~\cite{zampetti2017open}.

@@

% TODO caro: maybe convert this into a proper ordered list?
As a consequence, we took it upon ourselves (step 2) to provide
implementations of the three most promising automatic build log analysis
techniques we observed in the literature survey:
a technique 1) able to automatically synthesize regular expressions,
called \emph{Program Synthesis by Example (PBE)}, 2) based on natural
language processing, doing \emph{Common Text Similarity (CTS)}	and 3)
inspired by information-retrieval employing \emph{Keyword Search (KWS)}.
Even though these three approaches to analyze
build logs have different strengths and weaknesses, the literature
survey shows that their use often seems arbitrary.
As a consequence,
in the second research question we ask:

@@

To answer this question, in the third step  of \Cref{fig:overview},
we conduct an empirical study assessing the performance
of these chunk retrieval techniques in
real-world projects under a variety of conditions.
Our study is based on the \emph{LogChunks} data
set~\cite{brandt2020logchunks}, which encompasses 797 build logs from
80 popular open-source projects and a broad range of different
build tools and programming languages.
Our results show that there is no technique that in general outperforms
the others.
However,
% TODO 1-2 sentences about results, inclduing some accuracy measures

@@

To guide the data extraction we created a template.
The template form contained questions to answer RQ 1.1 through RQ 1.3,
for example ``Is the technique explained in detail?'' and
``What is the source of the build logs?''.
As recommended by Petersen et
al.~\cite{petersen2015guidelines}, we started the mapping study with a
pilot study:
both authors extracted data from the same five articles and held a
consensus meeting to unify their understanding
of the data extraction template.
We divided the remaining 46 articles evenly between the two authors
and inspected them as closely as necessary to fill out the template
with confidence,
starting from occurrences of ``log'' or ``build'' within the article's
text.
For questions that did not require a boolean answer, we allowed assigning
multiple tags or categories as the first part (``preparation phase'')
of a virtual open card sorting~\cite{zimmermann2016card}.
An example for this is the source of the analyzed build logs,
where we assigned tags such as ``Travis CI'',
``Industrial'', ``Google'', or
``self built''.
Once we had classified all works, we went into the second phase
(``execution phase'') in which we grouped tags
together and found synonyms.
% the following adds to the example but is maybe unneeded / too much
For instance, in the case of the build log source we grouped
``Travis'' and ``travis ci'' together and assigned ``Industrial'' to
all articles that used build logs provided by a company.

@@

% TODO: this section is crucial.
% maybe we can highlight important findings, by categorizing them either
% as further subsubsections or making it bold in the text?
% TODO: find values for X and Y
Our literature survey shows that despite its relatively young age
(first occurrence of a paper on the topic in 2001, vast majority after
2016), \textbf{build log analysis is a widely established technique}
in the literature, with Y articles on it.

@@

Several articles noted the variety in log formats for different build
environments, which requires customizing build log techniques for
every supported build environment.
\Cref{fig:litsur:log_producer} shows that many studies and tools target
only a few distinct build environments (roughly, area A).
% TODO: passive ...
The majority of build environments was however only targeted by one
article each (B).
From this, we follow: (a) there is a nichè for custom parsers or even
specialized plugins integrated into the build process to make
post-factum
parsing of logs redundant.
Plugins  similar to, for example, the Maven plugin to report test
executions Bell et al.
%TODO cite
wrote, still have the disadvantage of not being able to work on historic
builds.
(b) the large tail of seldomly-investigated environments would benefit
from the existence of a more generic
solution.
% We implement build environment agnostic build log analysis
% techniques, which are customized to a specific project by
% the training examples which the user provides.
% In addition, the \emph{LogChunks}~\cite{brandt2020logchunks}
% data set we chose for our evaluation covers a broad range of build
% environments.
% % Further, the training examples also specify the log chunk targeted
% by the
% % retrieval, which enable our chunk retrieval techniques to retrieve
% % a broad range of information from build logs.

@@

% Moritz: The following is very obvious and could apply
% to any
% % technique.
% % kinda.
% It does give a descriptive error message on no match (and
% % that is not possible in other techniques)
% TODO Caro: Can you given an example of such an error message?
% % otherwise yes, nothing interesting happens in the application of pbe
% % though for consistency we might need this section? or merge
% config/application?
% Why does it return an error message? Shouldn't it also return
% <empty string>? Does it not rather return an error message if no
% program can be synthesized?
% TODO Moritz: a regular expression can match an empty string
% and that is different from not finding a match
% it returns error messages in both cases
% TODO: Ok, but what is the point of matching on an empty string
(if I understand what you mean)? you would not normally define such
an extraction?
% If this distinction is important it needs to be made more clear
\subsubsection{Application}
A run of PBE takes a build log file as input and applies the
synthesized regular expression program.
It then returns the substring
of the build log matched by the program or an error message if the
program found no match.

@@

\subsubsection{Configuration}
To configure chunk retrieval through text similarity we chose to use
the same concept of examples as for PBE
% Moritz: your question was: More details on this.
% How do you
% 'apply' VSM? On what?
% all the following sentences in this subsubsection describe our way
% of determining text similarity which as far as I understand and
% Annibale said is the VSM.
% word tokens, tf-idf, similarity measure
% I added a "as follows", think that is enough for explanation?
% Moritz: I think it is good, but the VSM is still unexplained.
%We need a
% (at least) 1-sentence introduction to it, at its first reference
% TODO Moritz: I removed all other references, this is now the first one.
% I hope I got your comment right and now the first and only reference
% is explained :)
and apply the Vector Space Model~\cite{schutze2008introduction}
as follows:
The lines of the output strings of the training examples define our
search query.
The algorithm splits the search query into single lines and
identifies tokens, in this case words.
Then we build a
document-term-frequency matrix over the lines from the search query
and prune very often or very rarely appearing words.
Finally, the
algorithm applies TF-IDF to the matrix, a best practice for natural
language queries~\cite{lee1997document}.

@@

When we train PBE with these two examples, it produces
a regular expression program similar to the one presented in
\Cref{lst:prose-program-simplified}.
The retrieval
starts after a colon and a line separator and before a capital
letter.
It ends before two line separators,
which is consistent with the two example
chunks from \Cref{lst:chunk-example}.
This shows, that the characters structuring the text around the chunk are
crucial for regular expressions to identify a log chunk.
In fact, if we add the third example from \Cref{lst:chunk-example-3},
which has a
different structural representation, to the training set,
PBE is not able to synthesize a program and returns
``\texttt{no program found}.''
We later express this difference in structural representation through
dividing log chunks into \emph{structural categories}.
The chunks from \Cref{lst:chunk-example}
are in the \emph{same structural category}, while the chunk from
\Cref{lst:chunk-example-3} is in a
\emph{different structural category}.
The results of the empirical comparison study on the chunk retrieval
techniques
show that the structural representation of the targeted chunk greatly
influences
the performance of the chunk retrieval techniques.

@@

\subsection{Tool Implementation}
For the comparison study, we implemented each of the chunk retrieval
techniques and a unifying interface.
The unified interface is written in Ruby and
calls the separate technique implementations over the command line.
We implemented PBE in C\# based on the Microsoft PROSE
library~\cite{prose2019webpage} and CTS, KWS, and RLR using
R and the text2vec library~\cite{text2vec2019webpage}.

@@

\subsection{Study Process}
% TODO: I think you need to introduce a concept that makes clear what one
% ``folder'' consists of ...
% TODO: What does this mean -- we begin by? Aren't we perhaps missing that
% we operate on all repos in LogChunks?
We begin by selecting all examples in \emph{LogChunks}
that correspond to one GitHub repository---and thus the
% TODO () is meant to reference the figure, but this the figure is not
%actually mentioned.
same build environment (1).
% TODO: What does this mean? Of what nature are the examples? How were
%they chosen?
% Isn't it rather the case that for each repository, we built traning
%ets of size 1, 2,
% ...
%5 build logs?
One to five examples make up the training set (2).
We vary its size to measure how many examples we need to
confidently configure a technique.
The training set is so small because the examples have to be
manually created by a user.
A larger training set would oppose our goal of providing techniques
that can be configured with little effort
(see \Cref{sec:lit-sur:discussion}).
% TODO: The example? why only one? And what is an example? An example
%build log?
% Then it would be good to say so.
The example chronologically after the training set makes up the
test set (3).
In this way, we train on examples from past build logs and test on
more recent logs.
We chose a test set size of 1, as users of the techniques
should receive reliable results on every build log they analyze.
% TODO and why is this important? Doesn't everyone want such stability
%of results?
% motivation why only one test example.
we can also leave that out
% and answer if we are asked about it
In the next step, we configure a chunk retrieval technique with
the examples from the training set (4) and apply the technique to the
training set (5), which produces the retrieval output (6).
Based on this, we collect the following measurements (7):
the size of the training set (\textbf{RQ2.1}),
the number of structural categories in the training
set (\textbf{RQ2.2}),
the lines of the retrieval output ($\mathit{RetrievedLines}$),
and the oracle: the output defined in the test example
($\mathit{DesiredLines}$).
From these we calculate accuracy metrics (\textbf{RQ2.3}):

@@

% TODO what about referencing the combined figure at first: something
%like shows the results of CTS on two beanplots of ...
%and ...., and then the subfigure
% TODO this goes for all techniques
% TODO QUESTION: yes, though then we have more redundant text
% that repeats everywhere and is already in the captions
% -> should we add it for the first time we include a type of figure?
\Cref{fig:recall-precision-examplecount-CTS} presents precision,
recall, and F$_{1}$-score of chunk retrieval using CTS for an
increasing number of training examples.
When using one to five
training examples, the size of the training set has no noticeable
influence on precision, recall or F$_{1}$-score of the chunk retrieval
with CTS.

@@

\noindent
\textbf{Retrieval Output Usage}
While KWS has the highest recall of all three techniques, its
precision is the lowest.
The output of a chunk retrieval with KWS is
well-suited to be read by humans, but ill-suited for consumption by
automated tools.

@@

% TODO we can't just have a paragraph in a scientific paper start with
Caveat, I think ;-) Also, we should say why we THINK this is not the case.
Otherwise we weaken our work unduly
\textbf{Caveat!}
This is a preliminary hypothesis based on the results
from our comparison study.
The recommendations could therefore be
influenced by idiosyncrasies of our specific implementation of the
chunk retrieval techniques as well as the \emph{LogChunks}
data set.

@@

>>>>

@@

When a build or one of its steps fails, developers typically scavenge
the logs for information related to the source of the error.  This
manual activity is tedious and prone to
error~\cite{santolucito2018statically}.  As a single build log can be
over 50 megabytes large~\cite{beller2017oops}, finding the small chunk
of information in it which is linked to the error, is akin to
searching for a needle in a haystack.  In addition to helping
understand and fix build failures, the information richness of build
logs enables a plethora of other applications: with the help of build
logs, we can better understand and categorize CI failures and compile
errors~\cite{islam2017insights,seo2014programmers}, differentiate
testing practices~\cite{orellana2017differences,vassallo2017a-tale},
train classifiers to do ``predictive CI'' on the build or test outcome
and
duration~\cite{ni2017cost,bisong2017built,haghighatkhah2018test,machalica2019predictive},
or investigate static analysis tools in CI
builds~\cite{zampetti2017open}.

@@

% TODO caro: maybe convert this into a proper ordered list?
As a consequence, we took it upon ourselves (step 2) to provide
implementations of the three most promising automatic build log analysis
techniques we observed in the literature survey:
a technique 1) able to automatically synthesize regular expressions,
called \emph{Program Synthesis by Example (PBE)}, 2) based on information retrieval, using \emph{Common Text Similarity (CTS)}	and 3)
inspired by the manual adhoc-search for key parts in the log, employing \emph{Keyword Search (KWS)}.
Even though these three approaches---regular expressions, text similarity, keyword search---
have different strengths and weaknesses, the literature
survey shows that their use often seems arbitrary.
As a consequence,
in the second research question we ask:

@@

To answer this question, as a third step,
we conduct an empirical study assessing the performance
of these chunk retrieval techniques in
real-world projects under a variety of conditions.
Our study is based on the \emph{LogChunks} data
set~\cite{brandt2020logchunks}, which encompasses 797 build logs from
80 popular open-source projects and a broad range of different
build tools and programming languages.
Our results show that there is no technique that in general outperforms
the others.
However,
% TODO 1-2 sentences about results, inclduing some accuracy measures

@@

To guide the data extraction we created a template.
The template form contained questions to answer RQ 1.1 through RQ 1.3,
for example ``Is the technique explained in detail?'' and
``What is the source of the build logs?''
As recommended by Petersen et
al.~\cite{petersen2015guidelines}, we started the mapping study with a
pilot study:
both authors extracted data from the same five articles and held a
consensus meeting to unify their understanding
of the data extraction template.
We divided the remaining 46 articles evenly between the two authors
and inspected them as closely as necessary to fill out the template
with confidence,
starting from occurrences of ``log'' or ``build'' within the article's
text.
For questions that did not require a boolean answer, we allowed assigning
multiple tags or categories as the first part (``preparation phase'')
of a virtual open card sorting~\cite{zimmermann2016card}.
An example for this is the source of the analyzed build logs,
where we assigned tags such as ``Travis CI'',
``Industrial'', ``Google'', or
``self built''.
Once we had classified all works, we went into the second phase
(``execution phase'') in which we grouped tags
together and found synonyms.
% the following adds to the example but is maybe unneeded / too much
For instance, in the case of the build log source we grouped
``Travis'' and ``travis ci'' together and assigned ``Industrial'' to
all articles that used build logs provided by a company.

@@

% TODO: this section is crucial.
% maybe we can highlight important findings, by categorizing them either
% as further subsubsections or making it bold in the text?
% TODO: find values for X and Y
Our literature survey shows that despite its relatively young age
(first mentioned in a paper in 2001, vast majority after
2016), \textbf{build log analysis is an established technique}
in the literature, with 61 articles making use of it.

@@

Several articles noted the variety in log formats for different build
environments, which requires customizing build log techniques for
every supported build environment.
\Cref{fig:litsur:log_producer} shows that many studies and tools target
only a few distinct build environments (roughly, area A).
% TODO: passive ...
The majority of build environments was however only targeted by one
article each (B).
From this, we follow: (a) there is a niche for custom parsers or even
specialized plugins integrated into the build process to make
post-factum
parsing of logs redundant.
Plugins  similar to, for example, the Maven plugin to report test
executions Bell et al.
%TODO cite
wrote, still have the disadvantage of not being able to work on historic
builds.
(b) the large tail of seldomly-investigated environments would benefit
from the existence of a more generic
solution.
% We implement build environment agnostic build log analysis
% techniques, which are customized to a specific project by
% the training examples which the user provides.
% In addition, the \emph{LogChunks}~\cite{brandt2020logchunks}
% data set we chose for our evaluation covers a broad range of build
% environments.
% % Further, the training examples also specify the log chunk targeted
% by the
% % retrieval, which enable our chunk retrieval techniques to retrieve
% % a broad range of information from build logs.

@@

% Moritz: The following is very obvious and could apply
% to any
% % technique.
% % kinda.
% It does give a descriptive error message on no match (and
% % that is not possible in other techniques)
% TODO Caro: Can you given an example of such an error message?
% % otherwise yes, nothing interesting happens in the application of pbe
% % though for consistency we might need this section? or merge
% config/application?
% Why does it return an error message? Shouldn't it also return
% <empty string>? Does it not rather return an error message if no
% program can be synthesized?
% TODO Moritz: a regular expression can match an empty string
% and that is different from not finding a match
% it returns error messages in both cases
% TODO: Ok, but what is the point of matching on an empty string?
(if I understand what you mean)? you would not normally define such
an extraction?
% If this distinction is important it needs to be made more clear
\subsubsection{Application}
A run of PBE takes a build log file as input and applies the
synthesized regular expression program.
It then returns the substring
of the build log matched by the program or an error message if the
program found no match.

@@

\subsubsection{Configuration}
To configure chunk retrieval through text similarity we chose to use
the same concept of examples as for PBE
% Moritz: your question was: More details on this.
% How do you
% 'apply' VSM? On what?
% all the following sentences in this subsubsection describe our way
% of determining text similarity which as far as I understand and
% Annibale said is the VSM.
% word tokens, tf-idf, similarity measure
% I added a "as follows", think that is enough for explanation?
% Moritz: I think it is good, but the VSM is still unexplained.
%We need a
% (at least) 1-sentence introduction to it, at its first reference
% TODO Moritz: I removed all other references, this is now the first one.
% I hope I got your comment right and now the first and only reference
% is explained :)
% TODO Slightly better, but I still have a problem with this -- what
% exactly is the vector
% space model here?
% At a minimum, the text should mention what the vector is and how the model is used
and apply the Vector Space Model~\cite{schutze2008introduction}
as follows:
The lines of the output strings of the training examples define our
search query.
The algorithm splits the search query into single lines and
identifies tokens, in this case words.
Then we build a
document-term-frequency matrix over the lines from the search query
and prune very often or very rarely appearing words.
Finally, the
algorithm applies TF-IDF to the matrix, a best practice for natural
language queries~\cite{lee1997document}.

@@

When we train PBE with these two examples, it produces
a regular expression program similar to the one presented in
\Cref{lst:prose-program-simplified}.
The retrieval
starts after a colon and a line separator and before a capital
letter.
It ends before two line separators,
which is consistent with the two example
chunks from \Cref{lst:chunk-example}.
This shows, that the characters structuring the text around the chunk are
crucial for regular expressions to identify a log chunk.
In fact, if we add the third example from \Cref{lst:chunk-example-3} (which has a
different structural representation) to the training set,
PBE is not able to synthesize a program and returns
``\texttt{no program found}.''
We later express this difference in structural representation through
dividing log chunks into \emph{structural categories}.
The chunks from \Cref{lst:chunk-example}
are in the \emph{same structural category}, while the chunk from
\Cref{lst:chunk-example-3} is in a
\emph{different structural category}.
The results of the empirical comparison study on the chunk retrieval
techniques
show that the structural representation of the targeted chunk greatly
influences
the performance of the chunk retrieval techniques.

@@

\subsection{Tool Implementation}
For the comparison study, we implemented each of the chunk retrieval
techniques and a unifying interface.
The unified interface is written in Ruby and
calls the separate technique implementations over the command line.
We implemented PBE in C\# based on the Microsoft PROSE
library~\cite{prose2019webpage} and CTS, KWS, and RLR using
R and the {\tt text2vec} library~\cite{text2vec2019webpage}.

@@

\subsection{Study Process}
% TODO: I think you need to introduce a concept that makes clear what one
% ``folder'' consists of ...
% TODO: What does this mean -- we begin by? Aren't we perhaps missing that
% we operate on all repos in LogChunks?
We begin by selecting all examples in \emph{LogChunks}
that correspond to one GitHub repository---and thus the
% TODO () is meant to reference the figure, but this the figure is not
%actually mentioned.
same build environment (1).
% TODO: What does this mean? Of what nature are the examples? How were
%they chosen?
% Isn't it rather the case that for each repository, we built traning
%ets of size 1, 2,
% ...
%5 build logs?
One to five examples make up the training set (2).
We vary its size to measure how many examples we need to
confidently configure a technique.
The training set is so small because the examples have to be
manually created by a user.
A larger training set would oppose our goal of providing techniques
that can be configured with little effort
(see \Cref{sec:lit-sur:discussion}).
% TODO: The example? why only one? And what is an example? An example
%build log?
% Then it would be good to say so.
The example chronologically after the training set makes up the
test set (3).
In this way, we train on examples from past build logs and test on
more recent logs.
We chose a test set size of 1, as users of the techniques
should receive reliable results on every build log they analyze.
% TODO and why is this important? Doesn't everyone want such stability
%of results?
% motivation why only one test example.
%we can also leave that out
% and answer if we are asked about it
In the next step, we configure a chunk retrieval technique with
the examples from the training set (4) and apply the technique to the
training set (5), which produces the retrieval output (6).
Based on this, we collect the following measurements (7):
the size of the training set (\textbf{RQ2.1}),
the number of structural categories in the training
set (\textbf{RQ2.2}),
the lines of the retrieval output ($\mathit{RetrievedLines}$),
and the oracle: the output defined in the test example
($\mathit{DesiredLines}$).
From these we calculate accuracy metrics (\textbf{RQ2.3}):

@@

% TODO what about referencing the combined figure at first: something
%like shows the results of CTS on two beanplots of ...
%and ...., and then the subfigure
% TODO this goes for all techniques
% TODO QUESTION: yes, though then we have more redundant text
% that repeats everywhere and is already in the captions
% -> should we add it for the first time we include a type of figure?
% TODO Yes, that sounds good. You might even say Figure 12-15 show two
% plots in the short text under the section heading. It just makes the
% text flow so much better.
\Cref{fig:recall-precision-examplecount-CTS} presents precision,
recall, and F$_{1}$-score of chunk retrieval using CTS for an
increasing number of training examples.
When using one to five
training examples, the size of the training set has no noticeable
influence on precision, recall or F$_{1}$-score of the chunk retrieval
with CTS.

@@

\noindent
\textbf{Retrieval Output Usage.}
While KWS has the highest recall of all three techniques, its
precision is the lowest.
The output of a chunk retrieval with KWS is
well-suited to be read by humans, but ill-suited for consumption by
automated tools.

@@

% TODO we can't just have a paragraph in a scientific paper start with
% Caveat, I think ;-) Also, we should say why we THINK this is not the case.
% Otherwise we weaken our work unduly
\textbf{Caveat!}
This is a preliminary hypothesis based on the results
from our comparison study.
The recommendations could therefore be
influenced by idiosyncrasies of our specific implementation of the
chunk retrieval techniques as well as the \emph{LogChunks}
data set.

@@

Finally, we gave an
outline to guide future research efforts in the area of build log
analysis. Important themes are
% TODO Caro complete :-)
